Forecasting the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd
Update paper.qmd
Interpreted results more clearly and organized the structure of results section.

## Average Polling Percentage by State
From @fig-1, we can see that in the swing states of Arizona, Georgia, and North Carolina, Trump holds a narrow lead. In contrast, Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When looking at the overall average across these states, Harris slightly edges out Trump, indicating her overall advantage in these swing areas.

## Polling Percentage Distribution
The box plot in @fig-2 demonstrates the range and consistency of support for both candidates across each state. Trump shows a wider range of support, reflecting variability, while Harris has relatively steady polling percentages across states, maintaining higher consistency. Harris’s support is less variable and generally centers around higher polling percentages compared to Trump’s.

## Difference in Average Polling Percentage
@fig-3 highlights the difference in average polling percentages by state, with positive values indicating Trump’s lead and negative values indicating Harris's. Trump’s lead in the three states (Arizona, Georgia, and North Carolina) is relatively narrow, with margins of less than 1 percent. In contrast, Harris’s lead in the states favoring her (Pennsylvania, Wisconsin, Nevada, and Michigan) ranges from 1 to 4 percent, indicating her stronger support in the states where she leads.

@@

```{r}
#| label: fig-2
#| fig-cap: "Polling Percentage Distribution"
#| echo: false
#| warning: false
#| message: false
ggplot(data %>% filter(candidate_name %in% c("Donald Trump", "Kamala Harris")), 
       aes(x = state, y = pct, fill = candidate_name)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Polling Percentage Distribution by State",
       x = "State",
       y = "Polling Percentage",
       fill = "Candidate") +
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}
#| label: fig-3
#| fig-cap: "Difference in Average Polling Percentage"
#| echo: false
#| warning: false
#| message: false
data_diff <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state) %>%
  summarize(diff_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE) - 
                      mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE), .groups = "drop")

@@

## Swing State Polling Prediction Model
@fig-4 shows the trend of support rate changes for each candidate before the upcoming election, as well as the model's predicted values for the next 20 days.We predicts that Donald Trump may win based on our model.
```{r}
#| label: fig-4
#| fig-cap: "Swing State Polling Prediction Model"
#| echo: false
#| warning: false
#| message: false

@@

# Discussion

>>>>

## Average Polling Percentage by State
From @fig-1, we can see that in the swing states of Arizona, Georgia, and North Carolina, Trump holds a narrow lead. In contrast, Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When looking at the overall average across these states, Harris slightly edges out Trump, indicating her overall advantage in these swing areas. In our assumed electoral map, this projects Kamala Harris over the edge with 276 electoral votes as opposed to Donald Trump's 262. This shows that our model predicts Kamala Harris to win the 2024 Presidential Election if she is able to maintain her lead over all of Pennsylvania, Wisconsin, Nevada, and Michigan.

@@

```
## Polling Percentage Distribution
The box plot in @fig-2 demonstrates the range and consistency of support for both candidates across each state. Trump shows a wider range of support, reflecting variability, while Harris has relatively steady polling percentages across states, maintaining higher consistency. Harris’s support is less variable and generally centers around higher polling percentages compared to Trump’s.

```{r}
#| label: fig-2
#| fig-cap: "Polling Percentage Distribution"
#| echo: false
#| warning: false
#| message: false
ggplot(data %>% filter(candidate_name %in% c("Donald Trump", "Kamala Harris")), 
       aes(x = state, y = pct, fill = candidate_name)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Polling Percentage Distribution by State",
       x = "State",
       y = "Polling Percentage",
       fill = "Candidate") +
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
## Difference in Average Polling Percentage
@fig-3 highlights the difference in average polling percentages by state, with positive values indicating Trump’s lead and negative values indicating Harris's. Trump’s lead in the three states (Pennsylvania, Wisconsin, Nevada, and Michigan) is relatively narrow, with margins of less than 1 percent. In contrast, Harris’s lead in the states favoring her (Pennsylvania, Wisconsin, Nevada, and Michigan) ranges from 1 to 4 percent, indicating her stronger support in the states where she leads. Consequently, this difference in average polling percentages show that Kamala Harris' lead of 276 electoral votes is supported by a reliable margin and Donald Trump's lead in Arizona, Georgia, and North Carolina is turbulent.

```{r}
#| label: fig-3
#| fig-cap: "Difference in Average Polling Percentage"
#| echo: false
#| warning: false
#| message: false
data_diff <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state) %>%
  summarize(diff_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE) - 
                      mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE), .groups = "drop")

@@

## Swing State Polling Prediction Model
However, @fig-4 shows the trend of support rate changes for each candidate before the upcoming election, as well as the model's predicted values for the days until the election. This support rate changes model exhibits a higher rate of growth in Donald Trump's share of support. Although Kamala Harris is in the lead, Donald Trump's growth of support rate is an element that could stifle Harris' lead and drop her below her required 270 electoral votes. 
```{r}
#| label: fig-4
#| fig-cap: "Swing State Polling Prediction Model"
#| echo: false
#| warning: false
#| message: false

@@

# Discussion
## Turbulent Margins
Based on our results from @fig-1 predicted that Kamala Harris to win the presidential election with 276 electoral votes. Furthermore, @fig-2 showed that her lead in the states of Pennsylvania, Wisconsin, Nevada, and Michigan are stronger than Donald Trump's support in Pennsylvania, Wisconsin, Nevada, and Michigan. From @fig-2, we can also observe that Kamala Harris' lead in Nevada seems to be the weakest. In the framework of the electoral college system, losing Nevada to Donald Trump would not change the results as Nevada holds 6 electoral votes, which means that Kamala Harris would still possess the necessary 270 to win the election. However, this is not the case if Kamala Harris loses support in any of the three other states of Pennsylvania, Wisconsin, and Michigan as this would place her beneath the 270 seat threshold. In fact, losing any one of these states would result in Donald Trump's victory. @fig-3 showed that the support rate of Donald Trump is increasing faster than that of Kamala Harris', threatening Harris' lead over Pennsylvania, Wisconsin, and Michigan.
